---
title: "TSFinal"
author: "Christian Ortiz"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries


```{r}
library( astsa )
library( sqldf )
library( stringr )
library( dynlm )
library( tidyverse )
library( lubridate )
library( tseries )
library( forecast )
```


## Load Data

```{r}
PPI <- read.csv("PPI.csv")

CPI <- read.csv("CPI.csv")

Electricity <- read.csv("Electricity.csv")

Unemployment <- read.csv("Unemployment.csv")

```


## Data Wrangling

```{r}
# To Date 

CPI$DATE <- as.Date(CPI$DATE, format = '%Y-%m-%d')

PPI$DATE <- as.Date(PPI$DATE, format = '%Y-%m-%d')

Electricity$DATE <- as.Date(Electricity$DATE, format = '%Y-%m-%d')

Unemployment$DATE <- as.Date(Unemployment$DATE, format = '%Y-%m-%d')


# Filter for 1970 Onwards

CPI <- filter(CPI, DATE > '1970-01-01')

PPI <- filter(PPI, DATE > '1970-01-01')

Electricity <- filter(Electricity, DATE > '1970-01-01')

Unemployment <- filter(Unemployment, DATE > '1970-01-01')

# Renaming

colnames(PPI)[2] <-  "PPI"

colnames(CPI)[2] <- "CPI"

colnames(Electricity)[2] <- "Electricity_Prices"

colnames(Unemployment)[2] <- "Unemployment_Rate"

# Change Electricity Type to Numeric

Electricity$Electricity_Prices <- as.numeric(Electricity$Electricity_Prices)

# Joining Data Frames by DATE

df <- inner_join(PPI, CPI,  by = "DATE")
df <- inner_join(df, Electricity, by = "DATE")
df <- inner_join(df, Unemployment, by = "DATE")

```



## EDA


```{r}
tsplot(df$CPI)
tsplot(df$PPI)
tsplot(df$Electricity)
tsplot(df$Unemployment)
```



```{r}
tsplot(diff(df$CPI))
tsplot(diff(df$PPI))
tsplot(diff(df$Electricity))
tsplot(diff(df$Unemployment))
```

Hmm. It seems that theres weird years for Electricity (t = ~85) & Unemployment (t = ~495 - likely COVID).
Other than the odd time points mentioned above, differencing would seem to make them stationary.

Now, for PPI and CPI, even after differencing, it seems that the variance condition of stationarity would be violated.

```{r}
tsplot(log(df$CPI))
tsplot(log(df$PPI))
tsplot(log(df$Electricity))
tsplot(log(df$Unemployment))
```
Differencing for Unemployment & Electricity seem to suffice to create visual Stationarity.
However, logging then differencing for CPI and PPI might be the better move.



```{r}
tsplot(diff(log(df$CPI)))
tsplot(diff(log(df$PPI)))
```
For above transformations of logging then differencing CPI and PPI, it seems that these transformations
bring the series closer to stationarity, but there are still a few points with a high degree of variation.


Numerically testing for stationarity post transformations below.

```{r}
adf.test(diff(df$Electricity))

adf.test(diff(df$Unemployment_Rate))

adf.test(diff(log(df$CPI)))

adf.test(diff(log(df$PPI)))

```

adf tests suggest that above transformations meet stationarity.

Creating dataframe with transformed variables below.

```{r}
transformed_CPI <- diff(log(df$CPI))
transformed_PPI <- diff(log(df$PPI))
transformed_Electricity <- diff(log(df$Electricity_Prices))
transformed_Unemploymnent <- diff(df$Unemployment_Rate)
DATE <- df$DATE[-1]

transformed_df <- as.data.frame(cbind(DATE,transformed_PPI, transformed_CPI, 
                                      transformed_Electricity, transformed_Unemploymnent))
colnames(transformed_df) <- colnames(df)
```

checking acf below.
```{r}
acf(transformed_df$PPI)
acf(transformed_df$CPI)
acf(transformed_df$Electricity_Prices)
acf(transformed_df$Unemployment_Rate)
```



## CPI Modelling


```{r}
cpi_electricity.lm <- lm(CPI ~ Electricity_Prices, data = transformed_df)
summary(cpi_electricity.lm)
plot(cpi_electricity.lm)
checkresiduals(cpi_electricity.lm)
```
There seem to be 2 influential observations that are likely influencing parameters heavily.
Residual diagnostic suggests that OLS and ACF assumptions for regression don't hold, correlated residuals. 
F-statistic (p-val = 0.488) & R-Squared =  0.005623 suggests that might not be a good fit. 


```{r}
cpi_unemployment.lm <- lm(CPI ~ Unemployment_Rate, data = transformed_df)
summary(cpi_unemployment.lm)
plot(cpi_unemployment.lm)
checkresiduals(cpi_unemployment.lm)
```

1 Influential Observation.
F-statistic:  9.02,  p-value: 0.002801
Residual diagnostic suggests that residuals might be correlated. 


## PPI Modelling


```{r}
ppi_electricity.lm <- lm(PPI ~ Electricity_Prices, data = transformed_df)
summary(ppi_electricity.lm)
plot(ppi_electricity.lm)
checkresiduals(ppi_electricity.lm)
```
Same influential observations as CPI ~ Electricity. 
F-statistic: 7.498,  p-value: 0.00639, suggests good fit. 
Not sure that assumptions hold here.


```{r}
ppi_unemployment.lm <- lm(PPI ~ Unemployment_Rate, data = transformed_df)
summary(ppi_unemployment.lm)
plot(ppi_unemployment.lm)
checkresiduals(ppi_unemployment.lm)
```

1 influential observation. Same as CPI ~ Unemployment. 
F-statistic: 33.56,  p-value: 1.209e-08, suggests good fit.
Not sure about how residual diagnostic holds up here.









