---
title: "FRED inflation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F,warning=F,fig.width=10)
```


## Necessary packages 

```{r}
library(quantmod) # Extracting data
library(tsibble)
library(tidyverse)
library(ggplot2)
library(scales)
library(aTSA)
library(dynlm)
library(car)
library(sandwich)
```

## Extracting the data from FRED:

This code chunk below uses the quantmod package to get FRED data via its API.

```{r cars}
cpi=getSymbols('CPIAUCSL',src='FRED',auto.assign=F)['1992-01-01/']
cpi=cpi/cpi[[1]]*100
ppi=getSymbols('PITGCG01USM661N',src='FRED',auto.assign=F)['1992-01-01/']
ppi=ppi/ppi[[1]]*100
energy=getSymbols('PNRGINDEXM',src='FRED',auto.assign=F)['1992-01-01/']
energy=energy/energy[[1]]*100
food=getSymbols('PFOODINDEXM',src='FRED',auto.assign=F)['1992-01-01/']
food=food/food[[1]]*100
used_cars=getSymbols('CUSR0000SETA02',src='FRED',auto.assign=F)['1992-01-01/']
used_cars=used_cars/used_cars[[1]]*100
# I think you should turn govt_spending into an index using Jan 1992 as base. That way, it fits with how everything else is also currently an index.

housing=getSymbols('CPIHOSNS',src='FRED',auto.assign=F)['1992-01-01/']
housing=housing/housing[[1]]*100
govt_spending=getSymbols('G160371A027NBEA',src='FRED',auto.assign=F)['1992-01-01/']
#govt_spending=govt_spending/govt_spending[[1]]*100
personal_savings_rate=getSymbols('PSAVERT',src='FRED',auto.assign=F)['1992-01-01/'] # in percent
#personal_savings_rate=personal_savings_rate/personal_savings_rate[[1]]*100
```

## Merging:

This code chunk merges the individual extracted xts series into one object
```{r pressure, echo=FALSE}
list_ts=list(cpi,ppi,energy,food,used_cars,govt_spending,personal_savings_rate,housing)
final_xts=do.call(merge,list_ts)
# Final object omitting all NA variables due to lack of historical data
final_xts=na.approx(final_xts)
final_xts=na.locf(final_xts)
final_ts_df=final_xts%>%fortify.zoo()%>%as_tsibble()
colnames(final_ts_df)=c("Date","CPI","PPI","Energy","Food","Used_cars","Govt_spending","Personal_savings","Housing_prices")
colnames(final_xts)=c("CPI","PPI","Energy","Food","Used_cars","Govt_spending","Personal_savings","Housing_prices")
```


## EDA:
For all indices
```{r}
ggplot(final_ts_df,aes(x=Date))+
  geom_line(aes(y=PPI,color="PPI"))+
  geom_line(aes(y=CPI,color="CPI"))+
  geom_line(aes(y=Energy,color="Global price of energy index"))+
  geom_line(aes(y=Food,color="Global price of Food index"))+
  geom_line(aes(y=Used_cars,color="Used car price index"))+
  geom_line(aes(y=Housing_prices,color="Housing price index"))+
  theme_bw()+ylab("Index value (base = 100)")+xlab(NULL)+
  scale_x_date(date_breaks="4 year")+labs(colour="Index series")+ggtitle("Index series from 1992-01-01")
```

```{r}
ggplot(data=NULL,aes(x=final_ts_df$Date[-1]))+
  geom_line(aes(y=diff(log(final_ts_df$PPI)),color="PPI"))+
  geom_line(aes(y=diff(log(final_ts_df$CPI)),color="CPI"))+
  geom_line(aes(y=diff(log(final_ts_df$Energy)),color="Global price of energy index"))+
  geom_line(aes(y=diff(log(final_ts_df$Food)),color="Global price of Food index"))+
  geom_line(aes(y=diff(log(final_ts_df$Used_cars)),color="Used car price index"))+
  geom_line(aes(y=diff(log(final_ts_df$Housing_prices)),color="Housing price index"))+
  theme_bw()+ylab("Percentage change")+xlab(NULL)+
  scale_x_date(date_breaks="4 year")+
  scale_y_continuous(labels=scales::percent,breaks=scales::pretty_breaks(n=20))+
  labs(colour="Index series")+ggtitle("Index series from 1992-01-01")
```

## Checking for stationarity and persistence after first differencing of log variables:

```{r,fig.height=10}
par(mfrow=c(3,3))
acf(diff(log(final_ts_df$Energy)),main="ACF for Energy")
acf(diff(log(final_ts_df$Food)),main="ACF for Food")
acf(diff(log(final_ts_df$Housing_prices)),main="ACF for Housing")
acf(diff(log(final_ts_df$Used_cars)),main="ACF for Used cars")
acf(diff(log(final_ts_df$PPI)),main="ACF for PPI")
acf(diff(log(final_ts_df$CPI)),main="ACF for CPI")
```

After first differencing, most variables still exhibit AR(1) autocorrelation, with the exception of housing, which exhibits cyclicality in its ACF.

## Model fitting
```{r}
final_ts_df['log_Energy']=log(final_ts_df$Energy)
final_ts_df['diff_log_Energy']=difference(final_ts_df$log_Energy)
final_ts_df['lag_diff_log_Energy']=lag(final_ts_df$diff_log_Energy)
final_ts_df['log_Food']=log(final_ts_df$Food)
final_ts_df['diff_log_Food']=difference(final_ts_df$log_Food)
final_ts_df['lag_diff_log_Food']=lag(final_ts_df$diff_log_Food)
final_ts_df['log_CPI']=log(final_ts_df$CPI)
final_ts_df['diff_log_CPI']=difference(final_ts_df$log_CPI)
final_ts_df['log_PPI']=log(final_ts_df$PPI)
final_ts_df['diff_log_PPI']=difference(final_ts_df$log_PPI)
single_lag_CPI_Food_Energy=lm(diff_log_CPI~diff_log_Energy+diff_log_Food,data=final_ts_df)
double_lag_CPI_Food_Energy=lm(diff_log_CPI~diff_log_Energy+lag_diff_log_Energy+diff_log_Food+lag_diff_log_Food,data=final_ts_df)
single_lag_PPI_Food_Energy=lm(diff_log_PPI~diff_log_Energy+diff_log_Food,data=final_ts_df)
double_lag_PPI_Food_Energy=lm(diff_log_PPI~diff_log_Energy+lag_diff_log_Energy+diff_log_Food+lag_diff_log_Food,data=final_ts_df)
```

Model Building - CPI vs Housing and Used Cars

```{r}
final_ts_df['log_Housing'] = log(final_ts_df$Housing_prices)
final_ts_df['diff_log_Housing'] = difference(final_ts_df$log_Housing)
final_ts_df['lag_diff_log_Housing'] = lag(final_ts_df$diff_log_Housing)
final_ts_df['log_Used_Cars'] = log(final_ts_df$Used_cars)
final_ts_df['diff_log_Used_Cars'] = difference(final_ts_df$log_Used_Cars)
final_ts_df['lag_diff_log_Used_Cars'] = lag(final_ts_df$diff_log_Used_Cars)

CPI_Hosuing_Used_Cars <- lm ( diff_log_CPI ~ diff_log_Housing + diff_log_Used_Cars , data = final_ts_df)
lag_CPI_Hosuing_Used_Cars <- lm ( diff_log_CPI ~ diff_log_Housing + lag_diff_log_Housing + diff_log_Used_Cars + lag_diff_log_Used_Cars, data = final_ts_df)

```

Model Evaluation - CPI vs Housing and Used Cars

```{r}

summary(CPI_Hosuing_Used_Cars)

plot(CPI_Hosuing_Used_Cars$residuals)

acf(CPI_Hosuing_Used_Cars$residuals)
```

1 lag
```{r}
summary(lag_CPI_Hosuing_Used_Cars)

plot(lag_CPI_Hosuing_Used_Cars$residuals)

acf(lag_CPI_Hosuing_Used_Cars$residuals)

linearHypothesis(lag_CPI_Hosuing_Used_Cars,c("lag_diff_log_Housing=0", "lag_diff_log_Used_Cars=0"),vcov=vcovHAC)

```

Model Building - PPi vs Housing and Used Cars

```{r}

PPI_Hosuing_Used_Cars <- lm ( diff_log_PPI ~ diff_log_Housing + diff_log_Used_Cars , data = final_ts_df)

lag_PPI_Hosuing_Used_Cars <- lm ( diff_log_PPI ~ diff_log_Housing + lag_diff_log_Housing + diff_log_Used_Cars + lag_diff_log_Used_Cars, data = final_ts_df)

```

Model Evaluation - PPI vs Housing and Used Cars
```{r}
summary(PPI_Hosuing_Used_Cars)

plot(PPI_Hosuing_Used_Cars$residuals)

acf(PPI_Hosuing_Used_Cars$residuals)

```

1 lag
```{r}

summary(lag_PPI_Hosuing_Used_Cars)

plot(lag_PPI_Hosuing_Used_Cars$residuals)

acf(lag_PPI_Hosuing_Used_Cars$residuals)

linearHypothesis(lag_PPI_Hosuing_Used_Cars,c("lag_diff_log_Housing=0", "lag_diff_log_Used_Cars=0"),vcov=vcovHAC)

```

## Model examination

To compare models using no lags for each variable against models using 1 lag for each variable, we look at three things:
1) Their residual plots
2) Their ACF
3) Whether a model where the coefficients for lagged variables are coerced to zero has statistically significant difference from the model using 1 lag for each variable

```{r}
par(mfrow=c(1,2))
plot(rstandard(single_lag_CPI_Food_Energy)~fitted(single_lag_CPI_Food_Energy),ylab="Standardized residual",xlab="Fitted values",main="Single lag")
plot(rstandard(double_lag_CPI_Food_Energy)~fitted(double_lag_CPI_Food_Energy),ylab="Standardized residual",xlab="Fitted values",
main="Double lag")
```

```{r}
par(mfrow=c(1,2))
acf(resid(single_lag_CPI_Food_Energy),main="Single lag")
acf(resid(double_lag_CPI_Food_Energy),main="Double lag")
```

```{r}
linearHypothesis(double_lag_CPI_Food_Energy,c("lag_diff_log_Energy=0","lag_diff_log_Food=0"),vcov=vcovHAC)
```

```{r,fig.width=10}
par(mfrow=c(1,2))
plot(rstandard(single_lag_PPI_Food_Energy)~fitted(single_lag_PPI_Food_Energy),ylab="Standardized residual",xlab="Fitted values",main="Single lag")
plot(rstandard(double_lag_PPI_Food_Energy)~fitted(double_lag_PPI_Food_Energy),ylab="Standardized residual",xlab="Fitted values",
main="Double lag")
```

```{r}
par(mfrow=c(1,2))
acf(resid(single_lag_PPI_Food_Energy),main="Single lag")
acf(resid(double_lag_PPI_Food_Energy),main="Double lag")
```

```{r}
linearHypothesis(double_lag_PPI_Food_Energy,c("lag_diff_log_Energy=0","lag_diff_log_Food=0"),vcov=vcovHAC)
```

For both CPI and PPI, we see that using up to the first lag of the first difference of the logged variable yields a better fit model than not using such lags.

## Model selection:

```{r}

```


