---
title: "FRED inflation"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F,warning=F,fig.width=10)
```


## Necessary packages 

```{r}
library(MASS) # for the stepAIC() function
library(quantmod) # Extracting data
library(tsibble)
library(tidyverse)
library(ggplot2)
library(scales)
library(aTSA)
library(dynlm)
library(car)
library(sandwich)
library(modelsummary)
library(leaps) # for the leaps() function
```

## Extracting the data from FRED and indexing:

This code chunk below uses the quantmod package to get FRED data via its API.

```{r cars}
cpi=getSymbols('CPIAUCSL',src='FRED',auto.assign=F)['1992-01-01/']
cpi=cpi/cpi[[1]]*100
ppi=getSymbols('PITGCG01USM661N',src='FRED',auto.assign=F)['1992-01-01/']
ppi=ppi/ppi[[1]]*100
energy=getSymbols('PNRGINDEXM',src='FRED',auto.assign=F)['1992-01-01/']
energy=energy/energy[[1]]*100
food=getSymbols('PFOODINDEXM',src='FRED',auto.assign=F)['1992-01-01/']
food=food/food[[1]]*100
used_cars=getSymbols('CUSR0000SETA02',src='FRED',auto.assign=F)['1992-01-01/']
used_cars=used_cars/used_cars[[1]]*100
housing=getSymbols('CPIHOSNS',src='FRED',auto.assign=F)['1992-01-01/']
Unemployment=getSymbols('UNRATE',src='FRED',auto.assign=F)['1992-01-01/']/100
Electricity_avg=getSymbols('APU000072610',src='FRED',auto.assign=F)['1992-01-01/']
Electricity_avg=Electricity_avg/Electricity_avg[[1]]*100
```

## Merging:

This code chunk merges the individual extracted xts series into one object
```{r pressure, echo=FALSE}
list_ts=list(cpi,ppi,energy,food,used_cars,housing, Unemployment, Electricity_avg)
final_xts=do.call(merge,list_ts)
# Final object omitting all NA variables due to lack of historical data
final_xts=na.approx(final_xts) # Linear approximation
final_xts=na.locf(final_xts)
final_ts_df=final_xts%>%fortify.zoo()%>%as_tsibble()
colnames(final_ts_df)=c("Date","CPI","PPI","Energy","Food","Used_cars","Housing_prices", "Unemployment_Rate", "AvgElectricityPrices")
colnames(final_xts)=c("CPI","PPI","Energy","Food","Used_cars","Housing_prices", "Unemployment_Rate", "AvgElectricityPrices")
```


## EDA:
For all indices
```{r,fig.width=10,echo=F}
ggplot(final_ts_df,aes(x=Date))+
  geom_line(aes(y=PPI,color="PPI"))+
  geom_line(aes(y=CPI,color="CPI"))+
  geom_line(aes(y=Energy,color="Global price of energy index"))+
  geom_line(aes(y=Food,color="Global price of Food index"))+
  geom_line(aes(y=Used_cars,color="Used car price index"))+
  geom_line(aes(y=Housing_prices,color="Housing price index"))+
  geom_line(aes(y=AvgElectricityPrices,color="Average Electricity Prices"))+
  theme_bw()+ylab("Index value (base = 100)")+xlab(NULL)+
  scale_x_date(date_breaks="4 year")+labs(colour="Index series")+ggtitle("Index series from 1992-01-01")+
  theme(text=element_text(size=12))
```

```{r,fig.width=10,echo=FALSE}
ggplot(final_ts_df,aes(x=Date))+
  geom_line(aes(y=Unemployment))+
  theme_bw()+ylab("Index value (base = 100)")+xlab(NULL)+
  scale_x_date(date_breaks="4 year")+labs(colour="Index series")+ggtitle("Index series of unemployment rate from 1992-01-01")+
  theme(text=element_text(size=12))
```
Big Spike for Unemployment during covid that might have a significant impact on model. This informs our rationale to test COVID as a factor of inflation.

All of our series are non-stationary in their current form. As such, we choose to:
1. Take the log of all indices (so this excludes unemployment rate)
2. Take the first difference of the logged series, and the first difference of unemployment rate.

```{r,fig.width=10}
ggplot(data=NULL,aes(x=final_ts_df$Date[-1]))+
  geom_line(aes(y=diff(log(final_ts_df$PPI)),color="PPI"))+
  geom_line(aes(y=diff(log(final_ts_df$CPI)),color="CPI"))+
  geom_line(aes(y=diff(log(final_ts_df$Energy)),color="Global price of energy index"))+
  geom_line(aes(y=diff(log(final_ts_df$Food)),color="Global price of Food index"))+
  geom_line(aes(y=diff(log(final_ts_df$Used_cars)),color="Used car price index"))+
  geom_line(aes(y=diff(log(final_ts_df$Housing_prices)),color="Housing price index"))+
  geom_line(aes(y=diff(log(final_ts_df$AvgElectricityPrices)),color="Average U.S City Electricity Prices"))+
  theme_bw()+ylab("Percentage change")+xlab(NULL)+
  scale_x_date(date_breaks="4 year")+
  scale_y_continuous(labels=scales::percent,breaks=scales::pretty_breaks(n=20))+
  labs(colour="First difference of log")+ggtitle("Index series from 1992-01-01")
```

```{r,fig.width=10}
ggplot(data=NULL,aes(x=final_ts_df$Date[-1]))+
  geom_line(aes(y=diff(final_ts_df$Unemployment_Rate),color="First difference of unemployment rate"))+ggtitle("First difference of unemployment rate from 1992-01-01")+theme_bw()+xlab(NULL)+scale_x_date(date_breaks="4 year")+
  scale_y_continuous(labels=scales::percent,breaks=scales::pretty_breaks(n=20))+labs(colour="First difference of log")+ylab("")
```

## Checking for stationarity and persistence of first difference:

```{r,fig.height=10}
par(mar = c(3,3,5,5))
par(mfrow=c(3,3))
acf(diff(log(final_ts_df$PPI)),main="ACF for PPI")
acf(diff(log(final_ts_df$CPI)),main="ACF for CPI")
acf(diff(log(final_ts_df$Energy)),main="ACF for Energy")
acf(diff(log(final_ts_df$Food)),main="ACF for Food")
acf(diff(log(final_ts_df$Housing_prices)),main="ACF for Housing")
acf(diff(log(final_ts_df$Used_cars)),main="ACF for Used cars")
acf(diff(final_ts_df$Unemployment_Rate),main="ACF for Unemployment Rate")
acf(diff(log(final_ts_df$AvgElectricityPrices)),main="ACF for Avg. U.S City Electricity Prices")
```


## Lag importance examination:
We want to run a regression of PPI and CPI against our response variables, but we also want to know if we should include any lags of our variables in the regressors as well. To do so, we choose to:
1. Run a regression of our variables with the lags with statistically significant autocorrelation from the ACF against CPI and PPI..
2. Use a model F-test to determine if a model without lags could fit as well as a model with lags.


Food and Energy

```{r}
final_ts_df['log_Energy']=log(final_ts_df$Energy)
final_ts_df['diff_log_Energy']=difference(final_ts_df$log_Energy)
final_ts_df['lag_diff_log_Energy']=lag(final_ts_df$diff_log_Energy)
final_ts_df['log_Food']=log(final_ts_df$Food)
final_ts_df['diff_log_Food']=difference(final_ts_df$log_Food)
final_ts_df['lag_diff_log_Food']=lag(final_ts_df$diff_log_Food)
final_ts_df['log_CPI']=log(final_ts_df$CPI)
final_ts_df['diff_log_CPI']=difference(final_ts_df$log_CPI)
final_ts_df['log_PPI']=log(final_ts_df$PPI)
final_ts_df['diff_log_PPI']=difference(final_ts_df$log_PPI)
```

Housing and Used Cars

```{r}
final_ts_df['log_Housing'] = log(final_ts_df$Housing_prices)
final_ts_df['diff_log_Housing'] = difference(final_ts_df$log_Housing)
final_ts_df['lag_12_diff_log_Housing'] = lag(final_ts_df$diff_log_Housing, k = 12)
final_ts_df['log_Used_Cars'] = log(final_ts_df$Used_cars)
final_ts_df['diff_log_Used_Cars'] = difference(final_ts_df$log_Used_Cars)
final_ts_df['lag_diff_log_Used_Cars'] = lag(final_ts_df$diff_log_Used_Cars)
```

Electricity and Unemployment
```{r}
final_ts_df['diff_Unemployment']=difference(final_ts_df$Unemployment_Rate)
final_ts_df['lag_diff_Unemployment'] = lag(final_ts_df['diff_Unemployment'])
final_ts_df['log_AvgElectricityPrices']=log(final_ts_df$AvgElectricityPrices)
final_ts_df['diff_log_AvgElectricityPrices']=difference(final_ts_df$log_AvgElectricityPrices)
final_ts_df['lag_4_diff_log_AvgElectricityPrices'] = lag(final_ts_df$diff_log_AvgElectricityPrices, k = 4)
```


Model Building and Coefficient Evaluation 
Using p-value of 0.05 above (coefficient will be starred)
For CPI

```{r,fig.width=10}
lag_CPI_Energy=lm(diff_log_CPI~diff_log_Energy+lag_diff_log_Energy,data=final_ts_df)
lag_CPI_Food=lm(diff_log_CPI~diff_log_Food+lag_diff_log_Food,data=final_ts_df)
lag_CPI_Housing=lm(diff_log_CPI~diff_log_Housing + lag_12_diff_log_Housing,data=final_ts_df)
lag_CPI_Used_Cars=lm(diff_log_CPI~diff_log_Used_Cars + lag_diff_log_Used_Cars,data=final_ts_df)
lag_CPI_Unemployment=lm(diff_log_CPI~diff_Unemployment + lag_diff_Unemployment,data=final_ts_df)
lag_CPI_AvgElectricityPrices=lm(diff_log_CPI~diff_log_AvgElectricityPrices + lag_4_diff_log_AvgElectricityPrices, data = final_ts_df)
cpi_models=list(
  "Energy"=lag_CPI_Energy,
  "Food"=lag_CPI_Food,
  "Housing"=lag_CPI_Housing,
  "Used_cars"=lag_CPI_Used_Cars,
  "Unemployment"=lag_CPI_Unemployment,
  "Average Electricity Prices"=lag_CPI_AvgElectricityPrices
)
modelsummary(cpi_models,estimate="{estimate}{stars}",output = "table1.png")
```

For CPI, we determine that Energy and Food are significant up to their first lags based on their coefficients.


Model Building - PPI
For PPI

```{r,fig.width=10}
lag_PPI_Energy=lm(diff_log_PPI~diff_log_Energy+lag_diff_log_Energy,data=final_ts_df)
lag_PPI_Food=lm(diff_log_PPI~diff_log_Food+lag_diff_log_Food,data=final_ts_df)
lag_PPI_Housing=lm(diff_log_PPI~diff_log_Housing + lag_12_diff_log_Housing,data=final_ts_df)
lag_PPI_Used_Cars=lm(diff_log_PPI~diff_log_Used_Cars + lag_diff_log_Used_Cars,data=final_ts_df)
lag_PPI_Unemployment=lm(diff_log_PPI~diff_Unemployment + lag_diff_Unemployment,data=final_ts_df)
lag_PPI_AvgElectricityPrices=lm(diff_log_PPI~diff_log_AvgElectricityPrices + lag_4_diff_log_AvgElectricityPrices, data = final_ts_df)
ppi_models=list(
  "Energy"=lag_PPI_Energy,
  "Food"=lag_PPI_Food,
  "Housing"=lag_PPI_Housing,
  "Used_cars"=lag_PPI_Used_Cars,
  "Unemployment"=lag_PPI_Unemployment,
  "Average Electricity Prices"=lag_PPI_AvgElectricityPrices
)
modelsummary(ppi_models,estimate="{estimate}{stars}")
```

For PPI, we determine that Energy and Food are significant up to their first lags based on their coefficients. Additioanlly, unemployment rate is also significant for PPI up to the first lag.


## At this point, I will go ahead and insert Intervention term for COVID starting MARCH 2020

```{r}
# Something like this but for COVID
COVID <- c(seq(0, 0, length.out = 338), seq(1, 1, length.out = 21))
final_ts_df['COVID']=as.data.frame(COVID)
```

I will rerun the models having removed observations with NAs so that my model selection functions later down below can work. This will ultimately only remove 2 rows.

```{r}
# Here I will set the new final_ts_df object to only include non-NA having rows
final_ts_df <- final_ts_df[c(-1,-2),]
```


## Final model
We fit a model using all six variables with no lags, and the same model but including the first lags for Energy and Food Price as determined above.

### For CPI
```{r}
complete_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy  + diff_log_Food + diff_log_Housing +   diff_log_Used_Cars + diff_Unemployment + diff_log_AvgElectricityPrices, data = final_ts_df)
# We insert the first lag for Food and Energy as determined above.
complete_lag_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Food + lag_diff_log_Food + diff_log_Housing  +  diff_log_Used_Cars +  diff_Unemployment + diff_log_AvgElectricityPrices, data = final_ts_df)
```

```{r}
cpi_full_models=list(CPI_full=complete_fit_CPI,
     CPI_full_w_lags=complete_lag_fit_CPI)
modelsummary(cpi_full_models,estimate="{estimate}{stars}",,output="table2.png")
```


```{r,fig.height=10,fig.width=10}
par(mar = c(2,2,5,5))
par(mfrow=c(2,2))
plot(complete_fit_CPI,which=c(1))
plot(complete_lag_fit_CPI,which=c(1))
acf(complete_fit_CPI$residuals,main="Full CPI model")
acf(complete_lag_fit_CPI$residuals,main="Full CPI model with lags")
```

## For CPI w/ COVID Included

```{r}
COVID_complete_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy  + diff_log_Food + diff_log_Housing +   diff_log_Used_Cars + diff_Unemployment + diff_log_AvgElectricityPrices + factor(COVID), data = final_ts_df)

COVID_complete_lag_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Food + lag_diff_log_Food + diff_log_Housing  +  diff_log_Used_Cars +  diff_Unemployment + lag_diff_Unemployment + diff_log_AvgElectricityPrices + factor(COVID), data = final_ts_df)
```

```{r}
covid_cpi_models=list(
  "cpi_covid"=COVID_complete_fit_CPI,
  "cpi_covid_lag"=COVID_complete_lag_fit_CPI
)
modelsummary(covid_cpi_models,estimate="{estimate}{stars}",output="table4.png")
```

```{r,fig.height=10,fig.width=10}
par(mar = c(2,2,5,5))
par(mfrow=c(2,2))
plot(COVID_complete_fit_CPI,which=c(1))
plot(COVID_complete_lag_fit_CPI,which=c(1))
acf(COVID_complete_fit_CPI$residuals,main="Full CPI model with COVID")
acf(COVID_complete_lag_fit_CPI$residuals,main="Full CPI model with lags and COVID")
```

COVID doesn't seem to play a significant role for CPI model.

### For PPI
```{r}
complete_fit_PPI <- lm( diff_log_PPI ~ diff_log_Energy  + diff_log_Food + diff_log_Housing +   diff_log_Used_Cars + diff_Unemployment + diff_log_AvgElectricityPrices, data = final_ts_df)

complete_lag_fit_PPI <- lm( diff_log_PPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Food + lag_diff_log_Food + diff_log_Housing  +  diff_log_Used_Cars +  diff_Unemployment + diff_log_AvgElectricityPrices+lag_diff_Unemployment, data = final_ts_df)
```


```{r}
ppi_models=list(
  "ppi"=complete_fit_PPI,
  "ppi_lag"=complete_lag_fit_PPI
)
modelsummary(ppi_models,estimate="{estimate}{stars}",output="tablePPI.png")
```


```{r,fig.height=10,fig.width=10}
par(mar = c(2,2,5,5))
par(mfrow=c(2,2))
plot(complete_fit_PPI,which=c(1))
plot(complete_lag_fit_PPI,which=c(1))
acf(complete_fit_PPI$residuals,main="Full PPI model")
acf(complete_lag_fit_PPI$residuals,main="Full PPI model with lags")
```

```{r}
vif(complete_lag_fit_CPI)
vif(complete_lag_fit_PPI)
```

## For PPI w/ COVID Included

```{r}
COVID_complete_fit_PPI <- lm( diff_log_PPI ~ diff_log_Energy  + diff_log_Food + diff_log_Housing +   diff_log_Used_Cars + diff_Unemployment + diff_log_AvgElectricityPrices + factor(COVID), data = final_ts_df)

COVID_complete_lag_fit_PPI <- lm( diff_log_PPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Food + lag_diff_log_Food + diff_log_Housing  +  diff_log_Used_Cars +  diff_Unemployment+lag_diff_Unemployment + diff_log_AvgElectricityPrices + factor(COVID), data = final_ts_df)
```

```{r}
covid_ppi_models=list(
  "ppi_covid"=COVID_complete_fit_PPI,
  "ppi_covid_lag"=COVID_complete_lag_fit_PPI
)
modelsummary(covid_ppi_models,estimate="{estimate}{stars}",output="tablePPI2.png")
```


```{r,fig.height=10,fig.width=10}
par(mar = c(2,2,5,5))
par(mfrow=c(2,2))
plot(COVID_complete_fit_PPI,which=c(1))
plot(COVID_complete_lag_fit_PPI,which=c(1))
acf(COVID_complete_fit_PPI$residuals,main="Full PPI model with COVID")
acf(COVID_complete_lag_fit_PPI$residuals,main="Full PPI model with lags and COVID")
```

COVID does not play significant role in PPI model either. 

## Model Selection based on leaps() & stepAIC()

Since including COVID in the CPI or PPI model does not produce statistically significant coefficient estimates, I will go ahead and only evaluate the complete models without the COVID intervention term.

The Leaps below is based on Mallow's Cp which measures predictive ability of a fitted model. Good models have a small Mallow's Cp value  and a value close to the number of predictor variables in the model

### CPI Model Selection

```{r}

# Just included models commented out for my own purposes.

# complete_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy  + diff_log_Food + diff_log_Housing +   diff_log_Used_Cars + diff_Unemployment + diff_log_AvgElectricityPrices, data = final_ts_df)
# 
# complete_lag_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Food + lag_diff_log_Food + diff_log_Housing  +  diff_log_Used_Cars +  diff_Unemployment + lag_diff_Unemployment + diff_log_AvgElectricityPrices, data = final_ts_df)

complete_CPI.leaps.cp <- leaps(y = final_ts_df$diff_log_CPI, x = final_ts_df[,c(11,14,21,24,26,29)], method = c("Cp"))
complete_CPI.full.cp <- cbind(complete_CPI.leaps.cp$which, complete_CPI.leaps.cp$Cp)
complete_CPI.full.cp
```

Seems like Mallow's Cp values will differ slightly every time that the function runs, but values will lead to the same conclusions outlined below and thereafter.

From the results above, the model that produces a Mallow's Cp closest to the number of predictor variables in the complete_CPI model (6) is the model(1 0 1 1 1 1   6.664595) that excludes diff_log_food. However, going with the model that minimizes Mallow's Cp (1 0 1 1 0 1   4.686577) excludes diff_log_food & diff_Unemployment, which were both insignificant in the complete_fit_CPI model. So maybe we should see what a complete model without these predictor variables would look like? I'll go ahead and explore that later down below.

```{r}
complete_lag_CPI.leaps.cp <- leaps(y = final_ts_df$diff_log_CPI, x = final_ts_df[,c(11, 12, 14, 15, 21, 24,26, 27, 29)], method = c("Cp"))
complete_lag_CPI.full.cp <- cbind(complete_lag_CPI.leaps.cp$which, complete_lag_CPI.leaps.cp$Cp)
complete_lag_CPI.full.cp
```

From the results above, the model that produces a Mallow's Cp closest to the number of predictor variables in the complete_lag_fit_CPI model (9) is the model(1 1 0 1 1 1 1 1 1   9.179419) that excludes diff_log_food. However, going with the model that minimizes Mallow's Cp (1 1 0 1 1 1 1 0 1   7.255232) excludes diff_log_food & lag_diff_Unemployment, which were both insignificant in the complete_lag_fit_CPI model. 


Now down below, I will go ahead and run the same function but the metric will now be Adjusted R-Squared, so we ultimately want to choose a model based on the LARGEST Adjusted R-Squared (this selection will only really be valid if we're working off the basis that the OLS assumptions are justified btw).

```{r}
complete_CPI.leaps.r <- leaps(y = final_ts_df$diff_log_CPI, x = final_ts_df[,c(11,14,21,24,26,29)], method = c("adjr2"))
complete_CPI.full.r <- cbind(complete_CPI.leaps.r$which, complete_CPI.leaps.r$adjr2)
complete_CPI.full.r
```

So it seems that from the complete_fit_CPI model, based on getting the maximum Adjusted R-Squared value, the best model to go with is (1 1 1 1 0 1 0.43868042) which is the model that excludes diff_Unemployment. 

```{r}
complete_lag_CPI.leaps.r <- leaps(y = final_ts_df$diff_log_CPI, x = final_ts_df[,c(11, 12, 14, 15, 21, 24,26, 27, 29)], method = c("adjr2"))
complete_lag_CPI.full.r <- cbind(complete_lag_CPI.leaps.r$which, complete_lag_CPI.leaps.r$adjr2)
complete_lag_CPI.full.r
```

From the complete_lag_fit_CPI, the model that achieves the largest adjusted R-squared value would be (1 1 1 1 1 1 1 0 1 0.573446674) which exludes lag_diff_unemployment.


Now below I will go ahead and run the stepAIC() function on both CPI models to look at how a backtrace from the full model affects the AIC and then look at which models produce the best AIC.

```{r}
stepAIC(complete_fit_CPI, direction = "backward")
```

From the results above, it looks like excluding diff_Unemployment & diff_log_food from complete_fit_CPI produces the model with the lowest AIC value of AIC=-4304.84

```{r}
stepAIC(complete_lag_fit_CPI, direction = "backward")
```

Running the same function on the complete_lag_fit_CPI model tells us that excluding lag_diff_unemployment & diff_log_food would be the model with the lowest AIC of -4398.77

From Running a Leaps based on Mallow's Cp, Ajusted R-Squared, & the stepAIC function, it seems clear that we should remove any Food & Unemployment related predictor variables from the full model in order to get a model that maximizes how much variability can be explained by the data and a model that has the best predictive capability. I will go ahead and run what could be a "Final" CPI model based on the results above.

## "Final" CPI Model: w/ & w/o relevant lags

```{r}
final_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy  +  diff_log_Housing +   diff_log_Used_Cars +  diff_log_AvgElectricityPrices, data = final_ts_df)

final_lag_fit_CPI <- lm( diff_log_CPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Housing  +  diff_log_Used_Cars + diff_log_AvgElectricityPrices, data = final_ts_df)
```

```{r}
modelsummary(list(
  "final_CPI"=final_fit_CPI,
  "final_CPI_w_lags"=final_lag_fit_CPI),estimate="{estimate}{stars}",output="table5.png")
```


```{r,fig.height=10,fig.width=10}
par(mar = c(3,3,5,5))
par(mfrow=c(2,2))
plot(final_fit_CPI,which=c(1))
plot(final_lag_fit_CPI,which=c(1))
acf(final_fit_CPI$residuals,main="Model with lags")
acf(final_lag_fit_CPI$residuals,main="Model without lags")
```

From running what could be the "final" CPI models, including & excluding relevant lagged variables, it seems that the complete model without lags has a better AIC score with marginally better or similar other metric (Residual standard error, Adjusted R-Squared, etc.) values. However, the initial complete_lag_fit_CPI that includes food and unemployment related variables has a marginally better AIC, but I think this is simply due to inclusion of a higher number of predictor variables NOT the inclusion of important/significant predictor variables. Ultimately, I would argue that the above final CPI models would be preferable. 


## Now to do the exact same thing for the PPI Models
### PPI Model Selection

Just to make sure, the PPI models are exactly the same to the respective CPI models with the exception of the complete_lag_fit_PPI model. Whereas complete_lag_fit_CPI includes lag_diff_unemployment, complete_lag_fit_PPI does not. This will ultimately change the criteria for analyzing the results from the leaps() output slightly (as there is one less predictor variable..

```{r}



complete_PPI.leaps.cp <- leaps(y = final_ts_df$diff_log_PPI, x = final_ts_df[,c(11,14,21,24,26,27,29)], method = c("Cp"))
complete_PPI.full.cp <- cbind(complete_PPI.leaps.cp$which, complete_PPI.leaps.cp$Cp)
complete_PPI.full.cp
```

From the results above, the model that produces a Mallow's Cp closest to the number of predictor variables in the complete_PPI model (6) is the model(1 1 1 0 1 1   5.802378) that excludes diff_log_Used_Cars. However, going with the model that minimizes Mallow's Cp (1 1 1 0 1 0   3.822223) excludes diff_log_Used_Cars & diff_log_AvgElectricityPrices. 

```{r}
complete_lag_PPI.leaps.cp <- leaps(y = final_ts_df$diff_log_PPI, x = final_ts_df[,c(11, 12, 14, 15, 21, 24,26,27, 29)], method = c("Cp"))
complete_lag_PPI.full.cp <- cbind(complete_lag_PPI.leaps.cp$which, complete_lag_PPI.leaps.cp$Cp)
complete_lag_PPI.full.cp
```

From the results above, the model that produces a Mallow's Cp closest to the number of predictor variables in the complete_lag_fit_PPI model (8) is the model(1 1 1 1 1 0 1 1   7.538150) that excludes diff_log_Used_Cars. However, going with the model that minimizes Mallow's Cp (1 1 1 1 1 0 0 0   4.152958) excludes diff_log_Used_Cars, diff_Unemployment, & diff_log_AvgElectricityPrices.




```{r}
complete_PPI.leaps.r <- leaps(y = final_ts_df$diff_log_PPI, x = final_ts_df[,c(11,14,21,24,26,27,29)], method = c("adjr2"))
complete_PPI.full.r <- cbind(complete_PPI.leaps.r$which, complete_PPI.leaps.r$adjr2)
complete_PPI.full.r
```

So it seems that from the complete_fit_PPI model, based on getting the maximum Adjusted R-Squared value, the best model to go with is (1 1 1 0 1 0 0.54850982) which is the model that excludes diff_log_Used_Cars and diff_log_AvgElectricityPrices. 

```{r}
complete_lag_PPI.leaps.r <- leaps(y = final_ts_df$diff_log_PPI, x = final_ts_df[,c(11, 12, 14, 15, 21, 24,26,27, 29)], method = c("adjr2"))
complete_lag_PPI.full.r <- cbind(complete_lag_PPI.leaps.r$which, complete_lag_PPI.leaps.r$adjr2)
complete_lag_PPI.full.r
```

From the complete_lag_fit_PPI, the model that achieves the largest adjusted R-squared value would be (1 1 1 1 1 0 0 0 0.58475515) which excludes diff_log_Used_Cars, diff_Unemployment, diff_log_AvgElectricityPrices.

```{r}
stepAIC(complete_fit_PPI, direction = "backward")
```

From the results above, it looks like excluding diff_log_AvgElectricityPrices & diff_log_Used_Cars from complete_fit_PPI produces the model with the lowest AIC value of AIC=-3501.28

```{r}
stepAIC(complete_lag_fit_PPI, direction = "backward")
```

Running the same function on the complete_lag_fit_PPI model tells us that excluding diff_log_AvgElectricityPrices, diff_Unemployment & diff_log_Used_Cars would be the model with the lowest AIC of AIC=-3661.24


From the above analysis on PPI, it seems like it might be appropriate to exclude all predictor variables related to Used Cars, Unemployment, & Average electricity prices. With this in mind, I will go ahead and build what could be considered a "final" PPI model w/ & w/o relevant lags. 

## "Final" PPI Models

```{r}
final_fit_PPI <- lm( diff_log_PPI ~ diff_log_Energy  + diff_log_Food + diff_log_Housing, data = final_ts_df)

final_lag_fit_PPI <- lm( diff_log_PPI ~ diff_log_Energy + lag_diff_log_Energy + diff_log_Food + lag_diff_log_Food + diff_log_Housing+lag_diff_Unemployment, data = final_ts_df)
```

```{r}
modelsummary(list(
  "final_PPI"=final_fit_PPI,
  "final_PPI_w_lags"=final_lag_fit_PPI),estimate="{estimate}{stars}",output="table6.png")
```


```{r,fig.height=10,fig.width=10}
par(mar = c(2,2,5,5))
par(mfrow=c(2,2))
plot(final_fit_PPI,which=c(1))
plot(final_lag_fit_PPI,which=c(1))
acf(final_fit_PPI$residuals,main="Model without lags")
acf(final_lag_fit_PPI$residuals,main="Model with lags")
```
## Investigating if PPI and CPI affects PPI:
Ultimately, as everyday people, we are more concerned with CPI. As seen above, we have a model that parsimoniously describes some of the variables that can affect CPI. However, we want to know if PPI itself can affect CPI, by virtue of the producers passing on costs to consumers. Additionally, we want to see if CPI itself has autocrrelation.

```{r}
par(mar = c(3,3,5,5))
ccf(x=final_ts_df$diff_log_PPI,y=final_ts_df$diff_log_CPI,main="CPI cross correlation with PPI")
```

Based on our CCF plot, CPI inflation has significant positive correlation with PPI inflation at its zeroth and first lags, and has strong positive autocorrelation at a lag of 1
```{r}
final_ts_df['lag_diff_log_PPI']=lag(final_ts_df$diff_log_PPI,1)
final_lag_fit_full_CPI <- dynlm( diff_log_CPI ~ diff_log_PPI+lag_diff_log_PPI+diff_log_Energy + lag_diff_log_Energy + diff_log_Housing  +  diff_log_Used_Cars + diff_log_AvgElectricityPrices, data = final_ts_df)
```

```{r}
modelsummary(list(
  "final_model_without_PPI"=final_lag_fit_CPI,
  "final_model_with_PPI"=final_lag_fit_full_CPI
),estimate="{estimate}{stars}",output="table7.png")
```
```{r,fig.width=10}
par(mar = c(2,2,5,5))
par(mfrow=c(2,2))
plot(final_lag_fit_CPI,which=c(1))
plot(final_lag_fit_full_CPI,which=c(1))
acf(final_lag_fit_CPI$residuals,main="Model without PPI")
acf(final_lag_fit_full_CPI$residuals,main="Model with PPI")
```


## Summary:
We found that the energy prices currently and in the previous period, food prices currently and in the previous period, and housing prices are strong, positive correlated predictors for PPI. Energy prices and food prices make sense, given that they are common inputs for production processes. However, housing seems to be an anomalous predictor as there should be no reason why producers should care about housing prices, unless it affects their rent.

We found that the energy prices currently and in the previous period, housing prices, used cars prices, PPI currently and in the previous period are positively correlated predictors for CPI. This makes sense as these are items that factor into the current calculation for CPI, while PPI factors indirectly because producers, facing higher costs, will pass on some/all of the costs to consumers, leading to price raises for purchased goods. However, the negative correlation with average electricity prices seems anomalous, as the intuition is in the opposite direction: higher electricity prices should feed through all processes that utilize electricity, leading to price increases across the board.

Despite headlines of inflation being a result of COVID, we found no evidence of COVID being a variable that positively correlates with CPI and PPI.





